Precise predictions for same-sign W-boson scattering at the LHC

### 2018-03-09 Draft/conclusion.tex
minor edits to showers (shower section and conclusion)

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented, 
mainly focused on the electroweak production mechanism which involves the scattering of massive vector-bosons.
Until very recently, when the complete calculation became available for the NLO QCD corrections (order $\mathcal O (\alphas\alpha^6)$),
the so-called VBS approximation was the standard for this kind of simulations. For this reason, various theoretical predictions 
have been compared to the full computation, both in a typical VBS fiducial region and also in more inclusive phase-space.
We have precisely quantified the differences that arise for several physical observables, 
in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed.
%%%%and should be representative of the quality of the approximations for other VBS signatures.
Besides the study of fixed-order predictions, we have also investigated the impact of parton shower.
To that end, several LO and NLO event generators 
which are able to perform matching to parton showers have been employed, and various observables have been thoroughly compared.
While in general observables which are described at NLO accuracy show reasonable agreement among the tools, larger differences can 
appear for those observables described at a lower accuracy, such as those that involve the third jet.
In particular such differences are quite prominent in the central-rapidity region, and are the largest for those simulations which employ {\sc Pythia8}.
The effect
has been understood, and it can be partially mitigated by changing the recoil scheme of the parton shower. These findings make it worth 
to further investigate these issues not only in the theoretical community, but also by experimental collaborations, for example by 
measuring related observables for similar processes.

>>>>

In the present article, a detailed study of the process $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$ at the LHC has been presented, 
mainly focused on the electroweak production mechanism which involves the scattering of massive vector-bosons.
Until very recently, when the complete calculation became available for the NLO QCD corrections (order $\mathcal O (\alphas\alpha^6)$),
the so-called VBS approximation was the standard for this kind of simulations. For this reason, various theoretical predictions 
have been compared to the full computation, both in a typical VBS fiducial region and also in more inclusive phase-space.
We have precisely quantified the differences that arise for several physical observables, 
in particular for the di-jet invariant mass and the rapidity-separation of the leading two jets.
This is the first time that such an in-depth study is performed.
%%%%and should be representative of the quality of the approximations for other VBS signatures.
Besides the study of fixed-order predictions, we have also investigated the impact of parton showering.
To that end, several LO and NLO event generators 
which are able to perform matching to parton showers have been employed, and various observables have been thoroughly compared.
While in general observables which are described at NLO accuracy show reasonable agreement among the tools, larger differences can 
appear for those observables described at a lower accuracy, such as those that involve the third jet.
In particular such differences are quite prominent in the central-rapidity region, and are the largest for those simulations which employ {\sc Pythia8}.
The effect
has been understood, and it can be partially mitigated by changing the recoil scheme of {\sc Pythia8} to distribute momenta within initial-final colour connections. These findings make it worth 
to further investigate these issues not only in the theoretical community, but also by experimental collaborations, for example by 
measuring related observables for similar processes.

### 2018-03-09 Draft/shower.tex
minor edits to showers (shower section and conclusion)

We now discuss how different predictions compare when the matching to parton shower is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to show results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} and {\sc WHIZARD} are used for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). The same parton showers will be employed for the LO results of {\sc PHANTOM} and {\sc Pythia8} will also be employed for the LO results of {\sc WHIZARD}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the Binoth Les Houcehs Accord
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed.\ZM{Which matching scheme is used here in VBFNLO?}
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected. Multiple-parton interactions are disabled.

@@

We now present the results of predictions matched to parton shower.
The total rates within VBS cuts are displayed in Tables~\ref{tab:PSratesLO} and
\ref{tab:PSratesNLO}, at LO and NLO
accuracy respectively. For {\sc MG5\_aMC},
the numbers with $\Gamma_{\rm resc}$ are rescaled to
take into account the width effects described in the above paragraph. At NLO accuracy, for {\sc MG5\_aMC} + {\sc Pythia8} and {\sc VBFNLO}-{\sc Dipole}, we also quote
theoretical uncertainties.
For the former, we show both PDF and scale uncertainties,\footnote{A preliminary study on PDF uncertainties in VBS has appeared
in Ref.~\cite{Schwan:2017yy}.} obtained via exact reweighting~\cite{Frederix:2011ss} by varying independently the renormalisation and factorisation
scales by a factor of two around the central value, Eq.~(\ref{eq:defscale}) (nine-point variations).
For the latter, we show the
three-point scale uncertainties, obtained by considering correlated variations of the renormalisation and factorisation scales. Theory uncertainties should have very little dependence on the tool employed.
We observe that, once the width effect is taken into
account, total rates from different tools agree within some or few per cents, respectively at LO and NLO. Larger discrepancies, however, will appear for differential observables, which we are going to discuss in
the following. Theory uncertainties on the total rates are very small, regardless of whether scale variations are estimated with 
independent or correlated variations of the renormalisation and factorisation scales.
Concerning differential distributions, for each observable we will display results in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO}, \emph{i.e.}\ the VBS approximation with $s$-channel).
For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above. As for the table, we show theoretical uncertainties for the NLO+PS samples
obtained with {\sc VBFNLO} and {\sc MG5\_aMC}:
again, for the first the band corresponds to three-point variations, while for the second the darker (lighter) band corresponds to nine-point
scale variations (plus PDF uncertainties, linearly added).

>>>>

We now discuss how different predictions compare when the matching to parton shower is included. For such
a comparison we expect larger discrepancy than what we found at fixed-order, as a consequence of the different
matching schemes, parton shower employed, and of other details of the matching (such as the choice of the parton shower initial scale). Among
the codes capable of providing fixed-order results, presented before, {\sc MG5\_aMC}, the {\sc Powheg-Box}, and {\sc VBFNLO}
can also provide results at (N)LO+PS accuracy. For {\sc VBFNLO} matched to {\sc Herwig} and the {\sc Powheg-Box}, we
restrict ourselves to show results only in the VBS approximation,
{\emph i.e.}\ the $s$-channel contributions are neglected here. Besides,
also {\sc PHANTOM} and {\sc WHIZARD} are used for LO+PS results.\\
{\sc MG5\_aMC}, which
employs the {\sc MC@NLO}~\cite{Frixione:2002ik} matching procedure, will be used together with {\sc Pythia8}~\cite{Sjostrand:2014zea} (version 8.223)
and {\sc Herwig++}~\cite{Bahr:2008pv, Bellm:2013hwb} (version 2.7.1). The same parton showers will be employed for the LO results of {\sc PHANTOM} and {\sc Pythia8} will also be employed for the LO results of {\sc WHIZARD}. For the {\sc Powheg-Box}, the namesake
matching procedure is employed~\cite{Nason:2004rx,Frixione:2007vw}, together with {\sc Pythia8} (version 8.230). {\sc VBFNLO} serves as a matrix-element and phase-space provider
for the {\sc Matchbox} module~\cite{Platzer:2011bc} of {\sc
Herwig7}~\cite{Bellm:2015jjp,Bellm:2017bvx}, using an extended version of the Binoth Les Houches Accord
interface~\cite{Binoth:2010xt,Alioli:2013nda,Andersen:2014efa}. The {\sc Matchbox} module makes it
possible to choose between {\sc MC\-@NLO}-like and {\sc Powheg}-like
matching. As parton shower, both the default angular-ordered shower as
well as the dipole shower can be employed. We here use the subtractive (MC@NLO-type) matching to these parton shower algorithms.
Whenever {\sc Pythia8}\ is used, the Monash tune~\cite{Skands:2014pea} is selected. Multiple-parton interactions are disabled.

@@

We now present the results of predictions matched to parton showers.
The total rates within VBS cuts are displayed in Tables~\ref{tab:PSratesLO} and
\ref{tab:PSratesNLO}, at LO and NLO
accuracy respectively. For {\sc MG5\_aMC},
the numbers with $\Gamma_{\rm resc}$ are rescaled to
take into account the width effects described in the above paragraph. At NLO accuracy, for {\sc MG5\_aMC} + {\sc Pythia8} and {\sc VBFNLO}-{\sc Dipole}, we also quote
theoretical uncertainties.
For the former, we show both PDF and scale uncertainties,\footnote{A preliminary study on PDF uncertainties in VBS has appeared
in Ref.~\cite{Schwan:2017yy}.} obtained via exact reweighting~\cite{Frederix:2011ss} by varying independently the renormalisation and factorisation
scales by a factor of two around the central value, Eq.~(\ref{eq:defscale}) (nine-point variations).
For the latter, we show the
three-point scale uncertainties, obtained by considering correlated variations of the renormalisation and factorisation scales. Theory uncertainties should have very little dependence on the tool employed.
We observe that, once the width effect is taken into
account, total rates from different tools agree within some or few per cents, respectively at LO and NLO. Larger discrepancies, however, will appear for differential observables, which we are going to discuss in
the following. Theory uncertainties on the total rates are very small, regardless of whether scale variations are estimated with 
independent or correlated variations of the renormalisation and factorisation scales.
Concerning differential distributions, for each observable we will display results in two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO}, \emph{i.e.}\ the VBS approximation with $s$-channel).
For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above. As for the table, we show theoretical uncertainties for the NLO+PS samples
obtained with {\sc VBFNLO} and {\sc MG5\_aMC}:
again, for the first the band corresponds to three-point variations, while for the second the darker (lighter) band corresponds to nine-point
scale variations (plus PDF uncertainties, linearly added).
